Cancer Applications of Choropleth Maps, and the Potential of Cartograms and Alternative Map Displays

### 2019-12-18 Response/Response.Rmd

*Reply :*

*Changes in the text:*

*Comment 6: *

6. Please clearly state the primary aim of this paper and structure it accordingl=['
]
*Reply :*

@@

*Changes in the text:*

*Comment 7: *

7. Conclusions must discuss the key focus and conclusions of this paper.

*Reply 7:* 
Small changes have been made to the conclusion that make the focus of the paper and recommendations clearer.The conclusion has three parts (1) mapping practices, and recommendations, (2) staistics applied and the pros and cons, and (3) user interaction with online maps and the benefits of incorporating this.

*Changes in the text 7:*

### 2019-12-18 Response/Response.Rmd

This review paper presented an exploration of cancer mapping methods, focusing on variations on choropleths for online cancer atlases.

**Editor suggestion:**
**We would suggest that the aim of the manuscript could be phrased to highlight the current approaches and limitations of typical visualisation methods for cancer maps, along with outlining possible alternative methods and challenges in implementing these?**

We have restructed our original submission to provide a clear path from current approaches used, to the alternative approaches that should be explored. We highlight the limitations of using current display methods, to encourage the use of alternative display methods. We also present the challenges that may be faced when implementing the alternative designs.

The manuscript title has been changed:
"Cancer Applications of Choropleth Maps, and the Potential of Cartograms and Alternative Map Displays"

@@

*Also, within the bowel cancer atlas data, the reporting of percent of males aged 50-54 with bowel cancer in 2016 is incorrect; it should be the % of the ERP in each area that are males aged 50-54 years. *

We have changed the image used from the Bowel Cancer Atlas. To reflect this change the description of the statistic used in the bowel cancer atlas of Australia ahs been corrected:

Bowel Cancer Australia Atlas:
...shows the average Standardised Incidence Ratio of colorectal cancer for Australian males from 2006 to 2010 in Australia. It is published by Bowel Cancer Australia https://www.bowelcanceraustralia.org/.

The following section outlines comments made by the reviewer and contains our responses after the suggested changes were implemented.

@@

*Reviewer A: *
*Major*
*Comment 1: *
The Abstract should be more focused- perhaps give an example of how cartogram mapping has been applied to data visualisation. What are the benefits of this approach and suggestions for the future.

*Reply 1:* 
The advantages of alternative mapping displays, including cartograms, has been included in the abstract. It clearly includes the changes to the map display made by alternative visualisation methods, and how they benefit the map user as they build an understanding of the spatial distribution.

@@

*Changes in the text:*

Cancer atlases communicate cancer statistics over geographic domains, typically with a choropleth map. They subdivide these domains into administrative regions such as countries, states, or suburbs. When communicating human-related statistics, the choropleth has a disadvantage in that it draws attention to sparsely populated rural areas to the neglect of small inner city areas. The smaller geographic areas are important to consider if they are densely populated. Alternative map displays, such as a cartogram or a hexagon tile map, can shift the attention of map users from the large rural areas by decreasing their size on the map display. This means alternative displays can be more effective at accurately communicating spatial patterns across spatial areas, by highlighting areas of interest. It is recommended that alternative displays are included in cancer atlases. In addition, with the ease of today's technology, user interaction with the displays is encouraged. Users should also be able to interactively display different statistics, such as incidence rate or relative incidence.

@@

*Comment 2: *

2. The key message seems to get lost in the Abstract. The Conclusion is vague.

*Reply 2:* 
Changes to the abstract have been made. Additionally, the conclusion has also been edited to be more direct in recommending the use or consideration of the alternatives presented.

*Changes in the text:* 
Please see the response to *Comment 1*.

@@

*Comment 3: *

3. Entire paper needs substantial English language editing

*Reply 3:* 
Thank you for this point. Substantial time has been taken to improve the language used in this paper.

@@

*Comment 4: *

4. Please avoid the use of sentences which have been used throughout the paper where you start with a reference number: such as “[4] discusses the use of (Map displays for disease data, Line 15-16) They do not make any sense and must be modified.

*Reply:* 
The referencing system was changed to implement the springer-vancouver style. We apologise for the issues this caused when exporting to a Word Document, the text has been modified appropriately.

@@

*Comment 5: *

5. Please provide numbers for Figures and Tables

*Reply 5:* 
this begins the discussion, andNumbers for figures and tables have been corrected. This paper was produced using Rmarkdown, this introduced errors with numbering figures, sections and tables. All numbering has now been manually defined.

@@

*Comment 6: *

6. Please clearly state the primary aim of this paper and structure it accordingly

*Reply 6:* 
The primary aim of this paper is to present alternative map displays as viable visualisation methods for presenting spatial distributions in cancer atlases. A choropleth map is commonly used in practice, it is described and it's limitations are highlighted. The alternative map displays address the limitations of the choropleth map. The methods and the limitations of these alternative displays are discussed.

@@

*Comment 7: *

7. Conclusions must discuss the key focus and conclusions of this paper.

*Reply 7:* 
Small changes have been made to the conclusion that make the focus of the paper and recommendations clearer. The conclusion has three parts (1) mapping practices, and recommendations, (2) statistics applied and the pros and cons, and (3) user interaction with online maps and the benefits of incorporating this.

@@

*Minor Comments*

- Please check the use of past and present tense
- There are many minor grammatical errors that need to checked.
- Please edit the last section of the Introduction. Either number each section and then refer to them or find another way to appropriately refer to each section.

All minor comments were addressed in the process of substantial editing.

We thank you for your consideration of our changes.

### 2019-12-18 review_article.Rmd
Review article and responses finalised

Cancer atlases communicate cancer statistics over geographic domains, typically with a choropleth map. They subdivide these domains into administrative regions such as countries, states, or suburbs. When communicating human-related statistics, the choropleth has a disadvantage in that it draws attention to sparsely populated rural areas to the neglect of small inner city areas. The smaller geographic areas are important to consider if they are densely populated. Alternative map displays, such as a cartogram or a hexagon tile map, can be more effective at accurately communicating spatial patterns. It is recommended that more use of these alternative displays is made. In addition, with the ease of today's technology, user interaction with the displays is encouraged. Users should also be able to interactively display different statistics, such as incidence rate or relative incidence.

@@

```{r auscartograms, fig.cap = "Cartograms showing melanoma incidence in Australia: (a) contiguous, partially population transformed, (b) non-contiguous shape preserved, (c) Dorling, (d) hexagon tile map. The contiguous cartogram has expanded the highly populated areas while preserving the full shapes of rural areas. If it accurately sized areas by population, the country would be unrecognizable. The shape-preserved is unreadable due to the small area sizes. The Dorling cartogram presents all areas but many are difficult to compare. The hexagon tile map provides a reasonable spatial distribution despite having isolated hexagons in the outback areas.", message=FALSE, warning=FALSE, out.width='100%', cache=FALSE}

@@

library(cowplot)
ggdraw() +
  draw_plot(rasterGrob(aus1), 0, .55, 0.5, 0.50) +
  draw_plot(rasterGrob(aus2), 0.5, 0.55, 0.5, 0.50) +
  draw_plot(rasterGrob(aus3), 0.0, 0.1, 0.5, 0.55) +
  draw_plot(rasterGrob(aus4), 0.5, 0.1, 0.5, 0.55) +
  draw_plot(aus_legend, 0, 0, 1, .1) +
  draw_plot_label(c("a", "b", "c", "d"), 
    c(0, .5, 0, 0.5), 
    c(1, 1, .5, .5), size = 20)
```

@@

Figure 3<!--\ref{fig:auscartograms}--> shows four main types of cartograms displaying melanoma incidence on Statistical Areas at Level 3 [@abs2016] in Australia. The version of a contiguous cartogram (a) has expanded the highly populated areas while preserving the full shapes of rural areas. It has not fully resolved the population transformation of areas, and if it had accurately sized areas by population, the country would be unrecognizable. The shape-preserved cartogram is unreadable, and it has reduced all areas to tiny spots on the map. Zooming in on a high-resolution output shows it does preserve the areas. The Dorling cartogram has lost all geographic context. The hexagon tile map provides a reasonable spatial distribution despite having too much white-space in the outback areas.

@@

<!-- summary of mapping -->
This paper provides an overview of mapping practices as commonly used for cancer atlases and recommends new approaches, such as cartograms and hexagon tile maps that should be adopted going forward. The conventional approach is the choropleth map, and it is widely used. When there are small areas, as occurs in Australia where the population is concentrated on the coast, the information about cancer can be lost, and alternatives are needed. Making an inset can clarify congested regions but this breaks the viewers' attention because they need to shift focus from the map to the inset, and if there are many congested areas, many insets would be needed. The map alternatives, like cartograms and their variations, can be useful to allow the spatial distribution of cancer data to be digested.

<!-- statistics -->
Many different statistics are commonly used for cancer displays. The most basic is the incidence rate, which is easy to understand. It is common to see relative rates in many maps which measure how far a region is above or below the average. The purpose of using a relative rate is, perhaps the desire to pinpoint the areas that need attention because they have higher than expected rates. However, we lose the incidence rate information and thus interpretability. A region might be much higher than average, but it may not be close to a health concern, because all regions have a low incidence.

<!-- interaction -->
Interaction with maps is an important component of public atlases, and is easy to add with today's technology. A purpose is to provide access to more information than is possible to display in a single map, without overwhelming the viewer. Too many choices though can similarly overwhelm a viewer, and thus decisions do need to be made about content to provide, or hide, for accurate and comprehensive communication of information. Similarly, providing ways for users to interact with the display encourages engagement, and creative, efficient, elegant, interactive tools elicit curiosity about the data.

>>>>

Cancer atlases communicate cancer statistics over geographic domains, typically with a choropleth map. They subdivide these domains into administrative regions such as countries, states, or suburbs. When communicating human-related statistics, the choropleth has a disadvantage in that it draws attention to sparsely populated rural areas to the neglect of small inner city areas. The smaller geographic areas are important to consider if they are densely populated. Alternative map displays, such as a cartogram or a hexagon tile map, can shift the attention of map users from the large rural areas by decreasing their size on the map display. This means alternative displays can be more effective at accurately communicating spatial patterns across spatial areas, by highlighting areas of interest. It is recommended that alternative displays are included in cancer atlases. In addition, with the ease of today's technology, user interaction with the displays is encouraged. Users should also be able to interactively display different statistics, such as incidence rate or relative incidence.

@@

```{r auscartograms, fig.cap = "Cartograms showing melanoma incidence in Australia: (a) contiguous, partially population transformed, (b) non-contiguous shape preserved, (c) Dorling, (d) hexagon tile map. The contiguous cartogram has expanded the highly populated areas while preserving the full shapes of rural areas. If it accurately sized areas by population, the country would be unrecognizable. The shape-preserved is unreadable due to the small area sizes. The Dorling cartogram presents all areas but many are difficult to compare. The hexagon tile map provides a reasonable spatial distribution despite having isolated hexagons in the outback areas.", message=FALSE, warning=FALSE, out.width='100%', cache=FALSE, fig.width=8, fig.height=6, dpi=300}

@@

library(cowplot)
ggdraw(xlim = c(0,1), ylim = c(0,1.2)) +
  draw_plot(rasterGrob(aus1), 0, .65, 0.5, 0.50) +
  draw_plot(rasterGrob(aus2), 0.5, 0.65, 0.5, 0.50) +
  draw_plot(rasterGrob(aus3), 0.0, 0.1, 0.5, 0.55) +
  draw_plot(rasterGrob(aus4), 0.5, 0.1, 0.5, 0.55) +
  draw_plot(aus_legend, 0, 0, 1, .1) +
  draw_plot_label(c("a", "b", "c", "d"), 
    c(0, .5, 0, 0.5), 
    c(1.15, 1.15, .6, .6), size = 20)
```

@@

Figure 3<!--\ref{fig:auscartograms}--> shows four main types of cartograms displaying melanoma incidence on Statistical Areas at Level 3 [@abs2016] in Australia. The version of a contiguous cartogram (a) has expanded the highly populated areas while preserving the full shapes of rural areas. It has not fully resolved the population transformation of areas, and if it had accurately sized areas by population, the country would be unrecognizable. The shape-preserved cartogram is unreadable, and it has reduced all areas to tiny spots on the map. Zooming in on a high-resolution output shows it does preserve the areas. The Dorling cartogram and the hexagon tile map provide reasonable displays of the spatial distribution, despite having too much white-space in the outback areas.

@@

<!-- summary of mapping -->
This paper provides an overview of mapping practices as commonly used for cancer atlases and recommends new approaches, such as cartograms and hexagon tile maps that should be adopted going forward. The conventional approach is the choropleth map, and it is widely used. The choropleth map suffers when there are small geographic units, as occurs in Australia where the population is concentrated on the coast, the information about the burden of cancer on those communities can be hidden. Making an inset can clarify congested regions but this breaks the viewers' attention as they shift focus from the map to the inset, and if there are many congested areas, many insets would be needed. The map alternatives implement tradeoffs between the familiar shapes, and the importance of the geographic areas in the context of the areas. Given the population or a cancer statistic for each area, the geographic size or shape will change. Alternative displays allow the spatial distribution of cancer data to be digested by map users.

<!-- statistics -->
Many statistics are commonly used in cancer displays. The most basic is the incidence rate. It is common to see relative rates which measure how far a region is above or below the average. The purpose of using a relative rate is, perhaps the desire to pinpoint the areas that need attention because they have higher than expected rates. 
<!--However, we lose the incidence rate information and thus interpretability. --> A region might be much higher than average, but it may not be close to a health concern, because all regions have a low incidence. Supplementary materials can allow map users to recognise when this occurs.

<!-- interaction -->
Interaction with maps is an important component of public atlases, and is easy to add with today's technology. The purpose is to provide access to more information than is possible to display in a single map, without overwhelming the viewer. Too many choices can similarly overwhelm a viewer, and thus decisions do need to be made about content to provide for accurate and comprehensive communication of information. Similarly, providing ways for users to interact with the display encourages engagement, and creative, efficient, elegant, interactive tools elicit curiosity about the data.
